---
title: "Data Wrangling Nonsense in a Basic Multilevel Modeling Presentation"
author: "J DeVaney"
date: "12/12/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(ggplot2)
library(cowplot)
library(lme4)
library(lmerTest)
library(sjPlot)

mgmt <- read.csv("mgmt2.csv", header = T, stringsAsFactors = F)
```

## What will we discuss?

- The project(s) for which this data was collected

- My involvement with these projects

- Examples of reproducible data wrangling methods

- An approach to analyzing this data using multilevel modeling

## My role

- Statistical consultant

- Not my data

- Not my area of expertise

- Making this analysis process possible *and much easier* for others

# Looking at our data...

## Data: Source and Considerations

> - Collected across multiple years from employed students in the Department of Management and Marketing and their work supervisors
> - Self-reported or supervisor-reported scores for 20 surveys across 16 days for each individual
> - 588 x 973 (`r 588*973`) observations!... sort of.
> - So much missing data, *and* this data frame is a collection of item responses.

# What is our solution? </br> Scoring and reformatting. </br> </br> How? </br> With reproducible data wrangling methods!

# We know the structure, let's gather more basic information.

## For a "list" of the surveys:
```{r surveys}
non.id <- names(subset(mgmt, select = -Assessment.ID))

survey_plus <- non.id %>% 
  str_split("Day") %>% unlist() %>% 
  map_chr(str_extract, "^[:alpha:]+") %>% 
  unique() %>% na.omit() %>% as.character()

(surveys <- survey_plus %>% .[-(which(. == "Case"))])
```

## And the number of days:
```{r num_days}
(num_days <- mgmt %>%
  names() %>%
  str_extract("[:digit:]{1,}$") %>%
  as.numeric() %>%
  max(na.rm = T))
```

# Now we want to calculate survey scores with the responses, but what about items that need to be reverse scored?

##
```{r rev_score}
rev_score <- function(x = survey, y = questionN, n = maxLikert){
  for(i in 1:num_days){
    mgmt[, which(
      names(mgmt) == paste0(
        deparse(substitute(x)),
        deparse(substitute(y)),
        "Day",
        i))] <<- mgmt[, which(
                        names(mgmt) == paste0(
                          deparse(substitute(x)),
                          deparse(substitute(y)),
                          "Day",
                          i))] %>%
      map_dbl(~ (n + 1) - .x)
  }
  print(paste0("Item ",
               deparse(substitute(x)), deparse(substitute(y)),
               " has been altered."))
}

```
## Process made almost too simple...
```{r rev_2}
#from codebook

rev_score(Complain, 3, 5)
rev_score(Fatigue, 1, 5)
```

# We want to be able to operate across days within surveys, for each survey. 

# Multidimensional iteration means... nesting for loops! </br> </br> First, we need a format for this kind of operation.

## Large list arrangement
```{r unscored_list}
list_arrange <- function(df = mgmt){
  assign("data_list_unscored", list(), .GlobalEnv)
  for(i in 1:length(surveys)){
    assign(paste0(surveys[i], "_responses"),
           df %>% select(matches(paste0("^", surveys[i])))
    ) ->> data_list_unscored[[i]]
  }
}
list_arrange(mgmt)
names(data_list_unscored) <- surveys
```

## Viewing data frames within large list:
```{r}
#via custom function
unscored_survey_df <- function(i){
  View(data_list_unscored[[deparse(substitute(i))]])
}
scored_survey_df <- function(i){
  View(data_list_scored[[deparse(substitute(i))]])
}
```

- unscored_survey_df(CWB)
- scored_survey_df(OCB)

# Now let's score our surveys.

##
```{r}
assign("data_list_scored", list(), .GlobalEnv)
for(i in 1:length(surveys)){
  data_list_scored[[i]] <- data.frame(
    matrix(
      data = rep(NA, times = nrow(data_list_unscored[[i]])*num_days),
      ncol = num_days, nrow = nrow(data_list_unscored[[i]])))
  data_list_scored[[i]] <- data_list_scored[[i]] %>%
    setNames(., c(paste0(surveys[i], "Day", 1:num_days)))
}
for(i in 1:length(surveys)){
  for(j in 1:num_days){
    data_list_scored[[i]][,
      paste0(surveys[i], "Day", j)] <- data_list_unscored[[i]][,
        str_subset(names(data_list_unscored[[i]]),
          paste0("Day", j, "$"))] %>%
    as.data.frame() %>% apply(1, mean)
  }
}
names(data_list_scored) <- surveys
```

# While we can operate on our data in these separated frames, we should flip to a long-listed format.

##
```{r}
for (i in 1:length(surveys)){
  data_list_scored[[i]] <- data_list_scored[[i]] %>%
    tibble::rownames_to_column()
  colnames(data_list_scored[[i]])[1] <- "id"
  
  data_list_scored[[i]] <- data_list_scored[[i]] %>%
    select(id, contains(surveys[i])) %>%
    pivot_longer(
           cols = contains(paste0(surveys[i])),
           names_to = "time",
           values_to = surveys[i])
  
  data_list_scored[[i]]$time <- data_list_scored[[i]]$time %>%
    str_extract("[:digit:]{1,}") %>% as.numeric()
}
```


# Now we need to recombine everything into a single dataframe.

##
```{r}
df_yikes <- bind_cols(data_list_scored[1:length(surveys)])
(df <- bind_cols(select(df_yikes, id, time), df_yikes[surveys]))
```

## Where are we now?

- We have a massive data frame with 20 variables, excluding id and time.

- We don't want a massive data frame with 20 variables.

- We're going to select variables based on hypotheses of the study.

## Condensing the variable list

- Task Performance: Simply, how well one performs their work. Positively correlated with regulatory focus and OCBs (Koopman et al., 2017)

- Prevention Focus: Motivation driven by desire to minimize negative outcomes (Zhang, Cornwell, & Higgins, 2014)

- Organizational Citizenship Behavior (OCB): Actions beyond a defined role in an organization that benefit other members or the larger collective (Wendler, Liu, & Zettler, 2018)

- Let's not forget *time*

## What do we want to know?

- Prevention Focus and OCB are both expected to have positive relationships with Task Performance.

- Zhang et al. (2014) suggested that Prevention Focus led to repetition of behavior due to initial actions being perceived as the most effective strategy to succeed, even under changing conditions that lead to negative consequences for such behavior.

- It is predicted that OCB moderates the degree to which Prevention Focus predicts Task Performance, as a higher likelihood of behaviors that benefit one's environment would decrease the likelihood of negative behavioral repetition to interfere with performance.


##
```{r}
adf_miss <- df %>%
  select(id, time, TaskPerf, PrevFocus, OCB)
(adf <- adf_miss %>%
  filter(!is.na(TaskPerf) & !is.na(PrevFocus) & !is.na(OCB)))
```

## Descriptives for data subset of interest
```{r descriptives, echo=F, message=FALSE, warning=FALSE}
adf[, c(3,4,5)] %>% psych::describe(fast = T) %>% .[, c(2, 3, 4, 5, 6)] %>% knitr::kable()
```

# Let's Visualize

## Model Equation {.flexbox .vcenter}

$$TP_{ti} = \gamma_{00} + \gamma_{10}PF_{ti} + \gamma_{20}OCB_{ti} + \mu_{0i} + \mu_{1i}PF_{ti} + \mu_{2i}OCB_{ti} + \epsilon_{ti}$$

## Tidyverse Packages

- Lionel Henry and Hadley Wickham (2019). purrr: Functional
  Programming Tools. R package version 0.3.3.
  https://CRAN.R-project.org/package=purrr
- H. Wickham. ggplot2: Elegant Graphics for Data Analysis.
  Springer-Verlag New York, 2016.
- Hadley Wickham, Romain François, Lionel Henry and Kirill
  Müller (2019). dplyr: A Grammar of Data Manipulation. R
  package version 0.8.3.
  https://CRAN.R-project.org/package=dplyr
- Hadley Wickham (2019). stringr: Simple, Consistent Wrappers
  for Common String Operations. R package version 1.4.0.
  https://CRAN.R-project.org/package=stringr
- Hadley Wickham and Lionel Henry (2019). tidyr: Tidy Messy
  Data. R package version 1.0.0.
  https://CRAN.R-project.org/package=tidyr

## More Packages

- Douglas Bates, Martin Maechler, Ben Bolker, Steve Walker
  (2015). Fitting Linear Mixed-Effects Models Using lme4.
  Journal of Statistical Software, 67(1), 1-48.
  doi:10.18637/jss.v067.i01.
- Lüdecke D (2019). _sjPlot: Data Visualization for Statistics
  in Social Science_. doi: 10.5281/zenodo.1308157 (URL:
  https://doi.org/10.5281/zenodo.1308157), R package version
  2.7.2, <URL: https://CRAN.R-project.org/package=sjPlot>.
- Claus O. Wilke (2019). cowplot: Streamlined Plot Theme and
  Plot Annotations for 'ggplot2'. R package version 1.0.0.
  https://CRAN.R-project.org/package=cowplot
  
## References

- Koopmann, J., Johnson, R. E., Wang, M., Lanaj, K., Wang, G., & Shi, J. (2019). A self-regulation perspective on how and when regulatory focus differentially relates to citizenship behaviors. *Journal of Applied Psychology,* 104(5), 629–641.
- Wendler, K., Liu, J., & Zettler, I. (2018). Honesty-humility interacts with context perception in predicting task performance and organizational citizenship behavior.*Journal of Personnel Psychology,*17(4), 161–171.
- Zhang, S., Cornwell, J. F. M., & Higgins, E. T. (2014). Repeating the past: Prevention focus fotivates repetition, even for unethical decisions. *Psychological Science,* 25(1), 179–187.



