---
title: "Data Wrangling Nonsense in a Basic Multilevel Modeling Presentation"
author: "J DeVaney"
date: "12/12/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(vtable)
library(ggplot2)
library(lme4)
library(lmerTest)
library(sjPlot)

mgmt <- read.csv("mgmt2.csv", header = T, stringsAsFactors = F)
```

## What will we discuss?

- The project(s) for which this data was collected
- My involvement with these projects
- Examples of reproducible data wrangling methods
- An approach to analyzing this data using multilevel modeling

# Looking at our data...

## Data: Source and Considerations

> - Collected across multiple years from employed students in the Department of Management and Marketing and their work supervisors
> - Self-reported or supervisor-reported scores for 20 surveys across 16 days for each individual
> - 588 x 973 (`r 588*973`) observations!... sort of.
> - So much missing data, *and* this data frame is a collection of item responses.

# What is our solution? </br> Scoring and reformatting. </br> </br> How? </br> With reproducible data wrangling methods!

# We know the structure, let's gather more basic information.

## For a "list" of the surveys:
```{r surveys}
non.id <- names(subset(mgmt, select = -Assessment.ID))

survey_plus <- non.id %>% 
  str_split("Day") %>% unlist() %>% 
  map_chr(str_extract, "^[:alpha:]+") %>% 
  unique() %>% na.omit() %>% as.character()

(surveys <- survey_plus %>% .[-(which(. == "Case"))])
```

## And the number of days:
```{r num_days}
(num_days <- mgmt %>%
  names() %>%
  str_extract("[:digit:]{1,}$") %>%
  as.numeric() %>%
  max(na.rm = T))
```

# Now we want to calculate survey scores with the responses, but what about items that need to be reverse scored?

##
```{r rev_score}
rev_score <- function(x = survey, y = questionN, n = maxLikert){
  for(i in 1:num_days){
    mgmt[, which(
      names(mgmt) == paste0(
        deparse(substitute(x)),
        deparse(substitute(y)),
        "Day",
        i))] <<- mgmt[, which(
                        names(mgmt) == paste0(
                          deparse(substitute(x)),
                          deparse(substitute(y)),
                          "Day",
                          i))] %>%
      map_dbl(~ (n + 1) - .x)
  }
  print(paste0("Item ",
               deparse(substitute(x)), deparse(substitute(y)),
               " has been altered."))
}

```
## Process made almost too simple...
```{r rev_2}
#from codebook

rev_score(Complain, 3, 5)
rev_score(Fatigue, 1, 5)
```

# We want to be able to operate across days within surveys, for each survey. </br> Multidimensional iteration means... nesting for loops! </br> </br> First, we need a format for this kind of operation.

## Large list arrangement
```{r unscored_list}
list_arrange <- function(df = mgmt){
  assign("data_list_unscored", list(), .GlobalEnv)
  for(i in 1:length(surveys)){
    assign(paste0(surveys[i], "_responses"),
           df %>% select(matches(paste0("^", surveys[i])))
    ) ->> data_list_unscored[[i]]
  }
}
list_arrange(mgmt)
names(data_list_unscored) <- surveys
```

## Viewing data frames within large list:
```{r}
#via custom function
unscored_survey_df <- function(i){
  View(data_list_unscored[[deparse(substitute(i))]])
}
scored_survey_df <- function(i){
  View(data_list_scored[[deparse(substitute(i))]])
}
```

- unscored_survey_df(CWB)
- scored_survey_df(OCB)

# Now let's score our surveys.

##
```{r}
assign("data_list_scored", list(), .GlobalEnv)
for(i in 1:length(surveys)){
  data_list_scored[[i]] <- data.frame(
    matrix(
      data = rep(NA, times = nrow(data_list_unscored[[i]])*num_days),
      ncol = num_days, nrow = nrow(data_list_unscored[[i]])))
  data_list_scored[[i]] <- data_list_scored[[i]] %>%
    setNames(., c(paste0(surveys[i], "Day", 1:num_days)))
}
for(i in 1:length(surveys)){
  for(j in 1:num_days){
    data_list_scored[[i]][,
      paste0(surveys[i], "Day", j)] <- data_list_unscored[[i]][,
        str_subset(names(data_list_unscored[[i]]),
          paste0("Day", j, "$"))] %>%
    as.data.frame() %>% apply(1, mean)
  }
}
names(data_list_scored) <- surveys
```

# While we can operate on our data in these separated frames, we should flip to a long-listed format.

##
```{r}
for (i in 1:length(surveys)){
  data_list_scored[[i]] <- data_list_scored[[i]] %>% tibble::rownames_to_column()
  colnames(data_list_scored[[i]])[1] <- "id"
  
  data_list_scored[[i]] <- data_list_scored[[i]] %>%
    select(id, contains(surveys[i])) %>%
    pivot_longer(
           cols = contains(paste0(surveys[i])),
           names_to = "time",
           values_to = surveys[i])
  
  data_list_scored[[i]]$time <- data_list_scored[[i]]$time %>%
    str_extract("[:digit:]{1,}") %>% as.numeric()
}
```


# Now we need to recombine everything into a single dataframe.

##
```{r}
df_yikes <- bind_cols(data_list_scored[1:length(surveys)])
df <- bind_cols(select(df_yikes, id, time), df_yikes[surveys])
```

## Where are we now?

- We have a massive data frame with 20 variables, excluding id and time.
- We don't want a massive data frame with 20 variable.
- We're going to select variables based on hypotheses of the study.

##
```{r}
adf_miss <- df %>% select(id, time, TaskPerf, PrevFocus, OCB)
adf <- adf_miss %>% filter(!is.na(TaskPerf) & !is.na(PrevFocus) & !is.na(OCB))
```

# shit about the study variables

## Descriptives for data subset of interest

# visulization

# table

# model equation (reduced form?)

## discussion of results values


